---
title: "Lab1"
author: "Gabriel Dias Paiao"
date: "4/5/2020"
output: html_document
---

```{r setup, include=FALSE}
'https://github.com/toyetund-nmdp/teaching_APSC8280'

library(tidyverse)
library(lubridate)
library(ggplot2)

label = read.csv("dictionary.csv")
label = mutate(label, COUNTY_STATE = as.factor(COUNTY_CODE):as.factor(STATE_CODE))

data_2014 = read.csv("2014.csv")
data_2015 = read.csv("2015.csv")
data = rbind.data.frame(data_2014, data_2015)

pesticide=left_join(data, label)

```


## Problem 1 - Question 1
```{r, message = F}
sort(tapply(pesticide$HIGH_ESTIMATE, pesticide$COMPOUND, median, na.rm=T), decreasing = T)[1:10]

```

## Problem 1 - Question 2
```{r, message = F}
sort(tapply(pesticide$HIGH_ESTIMATE, pesticide$COMPOUND, sum, na.rm=T), decreasing = T)[1]

sort(tapply(pesticide$HIGH_ESTIMATE[pesticide$COMPOUND=="Glyphosate"], pesticide$STATE[pesticide$COMPOUND=="Glyphosate"], sum, na.rm=T), decreasing = T)[1]

sort(tapply(pesticide$HIGH_ESTIMATE[pesticide$COMPOUND=="Glyphosate"], pesticide$STATE[pesticide$COMPOUND=="Glyphosate"], sum, na.rm=T), decreasing = F)[1]

```

## Problem 1 - Question 3
```{r, message = F}
pest_low_est = drop_na(pesticide, LOW_ESTIMATE)

pest_low_est$perc_diff = ((pest_low_est$HIGH_ESTIMATE - pest_low_est$LOW_ESTIMATE) / ((pest_low_est$HIGH_ESTIMATE + pest_low_est$LOW_ESTIMATE) / 2)) *100

sort(tapply(pest_low_est$perc_diff, pest_low_est$STATE, mean, na.rm=T), decreasing = F)[1]

```

## Problem 1 - Question 4
```{r, message = F}
count_state = pesticide %>%
  group_by(STATE) %>%
  summarize(count = n_distinct(COMPOUND)) %>%
  arrange(desc(count))

print(count_state, n=1)

```

## Problem 1 - Question 5
```{r, message = FALSE}
amount_county = pesticide %>%
  group_by(COUNTY_STATE) %>%
  summarize(total = sum(HIGH_ESTIMATE)) %>%
  arrange(desc(total))

right_join(label, amount_county)[1,1:6]
```

## Problem 1 - Question 6
```{r, message = FALSE}
amount_county = pesticide %>%
  group_by(STATE) %>%
  summarize(total = n_distinct(COUNTY)) %>%
  arrange(desc(total))

print(amount_county, n=1)
```

## Problem 1 - Question 7
```{r, message = FALSE}
amount_by_state_2014 = pesticide %>%
  filter(YEAR == 2014) %>%
  group_by(STATE, add = T) %>%
  summarize(amount = median(HIGH_ESTIMATE), n_counties = n_distinct(COUNTY)) %>%
  mutate(usage_per_county = amount / n_counties) %>%
  mutate(YEAR = 2014) %>%
  arrange(desc(usage_per_county))
  
amount_by_state_2014[1:10,-c(2,3)]


amount_by_state_2015 = pesticide %>%
  filter(YEAR == 2015) %>%
  group_by(STATE, add = T) %>%
  summarize(amount = median(HIGH_ESTIMATE), n_counties = n_distinct(COUNTY)) %>%
  mutate(usage_per_county = amount / n_counties) %>%
  mutate(YEAR = 2015) %>%
  arrange(desc(usage_per_county))
  
amount_by_state_2015[1:10,-c(2,3)]

```


## Problem 1 - Question 8
```{r, message = FALSE}
amount_2014 = filter(pesticide, YEAR=="2014") %>%
  group_by(STATE) %>%
  summarize(total = n_distinct(COMPOUND)) %>%
  arrange(desc(total))

print(amount_2014, n=1)

amount_2015 = filter(pesticide, YEAR=="2015") %>%
  group_by(STATE) %>%
  summarize(total = n_distinct(COMPOUND)) %>%
  arrange(desc(total))

print(amount_2015, n=1)

```

## Problem 1 - Question 9
```{r, message = FALSE}
top_pesticide = pesticide %>%
  group_by(STATE) %>%
  summarize(total = median(HIGH_ESTIMATE)) %>%
  arrange(desc(total))

print(top_pesticide, n=5)

```

## Problem 1 - Question 10
```{r, message = FALSE}
most_states = pesticide %>%
  group_by(COMPOUND) %>%
  summarize(total = n_distinct(STATE)) %>%
  arrange(desc(total))

print(most_states, n=50)

```

## Problem 1 - Question 11
```{r, message = FALSE}
pesticide_state14 = filter(pesticide, YEAR =="2014") %>%
  group_by(COMPOUND) %>%
  summarize(total_2014 = n_distinct(STATE)) %>%
  arrange(desc(total_2014))

pesticide_state15 = filter(pesticide, YEAR =="2015") %>%
  group_by(COMPOUND) %>%
  summarize(total_2015 = n_distinct(STATE)) %>%
  arrange(desc(total_2015))

state_14_15 = mutate(left_join(pesticide_state14, pesticide_state15), pct_increase = ((total_2015-total_2014)/total_2014)*100)

print(arrange(state_14_15, desc(pct_increase)), n=1)

```

## Problem 1 - Question 12
```{r, message = FALSE}
use_state14 = filter(pesticide, YEAR =="2014") %>%
  group_by(STATE) %>%
  summarize(total_2014 = median(HIGH_ESTIMATE)) %>%
  arrange(desc(total_2014))

use_state15 = filter(pesticide, YEAR =="2015") %>%
  group_by(STATE) %>%
  summarize(total_2015 = median(HIGH_ESTIMATE)) %>%
  arrange(desc(total_2015))

state_14_15 = mutate(left_join(use_state14, use_state15), pct_increase = ((total_2015-total_2014)/total_2014)*100)

print(arrange(state_14_15, desc(pct_increase)), n=5)

```

## Problem 1 - Question 13 - We can see that the datasets for 2014 and 2015 do not comprehend the same range of estimates.
```{r, message = FALSE}

ggplot(data = pesticide) +
  geom_smooth(mapping = aes(y = HIGH_ESTIMATE, x = LOW_ESTIMATE)) +
  facet_wrap(facet = "YEAR")
  
rm(list = ls())
```


```{r, message = FALSE, include = FALSE}
accidents = read.csv("Accidents0514.csv")
vehicles = read.csv("Vehicles0514.csv")
casualties = read.csv("Casualties0514.csv")
head(accidents); dim(accidents)
head(casualties); dim(casualties)
head(vehicles); dim(vehicles)

```


## Problem 2 - Question 1
```{r, message = FALSE}
prop_acc = accidents %>%
  count(Urban_or_Rural_Area)

prop_acc = mutate(prop_acc, prop = n / sum(n)); prop_acc

```

## Problem 2 - Question 2 - There does not seem to exist a relationship between speed limit and proportion of accidents.
```{r, message = FALSE}
speed = accidents %>%
  group_by(Speed_limit) %>%
  summarize(total = sum(Number_of_Casualties))

speed = mutate(speed, prop = total / sum(total))

cor(speed$Speed_limit, speed$prop)
plot(speed$Speed_limit, speed$prop)
```

## Problem 2 - Question 3
```{r, message = FALSE}
prop_hour = accidents %>%
  mutate(hours = hm(Time, quiet = T)) %>%
  mutate(hr = hour(hours)) %>%
  left_join(casualties) %>%
  group_by(hr) %>%
  group_by(Casualty_Severity, add = TRUE) %>%
  summarize(total_fat = sum(Number_of_Casualties)) %>%
  group_by(hr) %>%
  mutate(prop_fat = total_fat / sum(total_fat))

prop_hour = filter(prop_hour, Casualty_Severity == 1)

plot(prop_hour$prop_fat~prop_hour$hr)

```

## Problem 2 - Question 4
```{r, message = FALSE}
trend_year = accidents %>%
  separate(Date, c("day", "month", "year")) %>%
  group_by(year) %>%
  count()

lm(trend_year$n~as.numeric(trend_year$year))$coefficients[2]

```

## Problem 2 - Question 5
```{r, message = FALSE}
male_female = vehicles %>%
  filter(Vehicle_Type == 9) %>%
  left_join(casualties) %>%
  filter(Sex_of_Driver %in% c(1,2)) %>%
  left_join(accidents) %>%
  group_by(Sex_of_Driver) %>%
  group_by(Accident_Severity, add = TRUE) %>%
  count() %>%
  group_by(Sex_of_Driver) %>%
  mutate(prop_fat = n / sum(n)) %>%
  filter(Accident_Severity == 1)

prop_m_f = male_female$prop_fat[1]/male_female$prop_fat[2]; prop_m_f
  
```

## Problem 2 - Question 6
```{r, message = FALSE}
decay = vehicles %>%
  filter(Age_of_Driver >= 17) %>%
  left_join(accidents) %>%
  group_by(Age_of_Driver) %>%
  count()

m1 = lm(log(decay$n)~decay$Age_of_Driver); m1$coefficients[2]

ggplot(decay, aes(x = decay$Age_of_Driver, y = decay$n)) +
  geom_point() +
  geom_line(data = fortify(m1), aes(x = decay$Age_of_Driver, y = exp(fitted(m1))))

```

## Problem 2 - Question 7
```{r, message = FALSE}
count_nice_soj = vehicles %>%
  left_join(accidents) %>%
  filter(Skidding_and_Overturning %in% c(1,2,3,4,5)) %>%
  filter(Weather_Conditions %in% c(1)) %>%
  count()

count_nice_none = vehicles %>%
  left_join(accidents) %>%
  filter(Skidding_and_Overturning %in% c(0)) %>%
  filter(Weather_Conditions %in% c(1)) %>%
  count()

count_rs_none = vehicles %>%
  left_join(accidents) %>%
  filter(Skidding_and_Overturning %in% c(0)) %>%
  filter(Weather_Conditions %in% c(2,3,5,6)) %>%
  count()

count_rs_soj = vehicles %>%
  left_join(accidents) %>%
  filter(Skidding_and_Overturning %in% c(1,2,3,4,5)) %>%
  filter(Weather_Conditions %in% c(2,3,5,6)) %>%
  count()

ratio_nice_weather = count_nice_soj / (count_nice_soj + count_nice_none)
ratio_rs_weather = count_rs_soj / (count_rs_soj + count_rs_none)

likelihood = ratio_rs_weather/ratio_nice_weather; likelihood

```